---
############################################################################################
#This code will run DEseq QC and differential analysis on the counts file, from rsubread   #
#This is work in progress and is currently only set up for my experiment                   #
#I plan to make all parts of the code relating to my experiment variable so that they can  #
#be set	for any experiment								   #
############################################################################################

---
title: "DEseq2_run"
author: "Steve W."
date: "December 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Get the input files from command line
```{r}
myfile <-commandArgs(trailingOnly=T)[1] #counts file
myfile2 <- commandArgs(trailingOnly=T)[2] #Sample sheet
peakfile <- commandArgs(trailingOnly=T)[3] # Master peak file (For gene annotation)
genetocheck <-commandArgs(trailingOnly=T)[3]# Gene to output differential analysis for
```

##Load in data files, count file, sample sheet, and original bed file, also get annotation programs loaded
```{r include=FALSE, echo=TRUE}
library(ChIPseeker)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
#myfile <- "/ifs/projects/proj078/1_atac_seq/SteveW_Rproject/Peak_counts.csv"  Use these to override command line arguments
#myfile2 <- "/ifs/projects/proj078/1_atac_seq/SteveW_Rproject/sample_sheet.csv"
#peakfile <- "/ifs/projects/proj078/1_atac_seq/4_data_for_steve/peakcalling_Jul18/mm_ATACCT8LightPulse30v1v2_R1.macs2_peaks.narrowPeak"
#genetocheck <- "fos"
counts <- as.matrix(read.csv(myfile,sep=",",row.names="Peaks"))
colnames <- read.csv(myfile2,sep=",", row.names=1)
peak <- readPeakFile(peakfile)
```

##Peak Annotation
```{r include=FALSE, echo=TRUE}
peakAnno <- annotatePeak(peakfile, tssRegion=c(-3000, 3000),
                         TxDb=txdb, annoDb="org.Mm.eg.db")
anno_table <- as.data.frame(peakAnno)
colnames(anno_table)[6] <- "peak"
```

##Set up experimental parameters to analyse
```{r include=FALSE, echo=TRUE}
library("DESeq2")
analysis <- DESeqDataSetFromMatrix(countData = counts,
                              colData = colnames,
                              design = ~ Light_Pulse + CT) #From sample sheet
```

##Pre filter the data
```{r message=FALSE}
keep <- rowSums(counts(analysis)) >= 10
analysis <- analysis[keep,]
```

##Carry out Differential Expression Analysis
```{r include=FALSE, echo=TRUE}
analysis <- DESeq(analysis)
res <- results(analysis)
resultsNames(analysis)
resLFC <- lfcShrink(analysis, coef=2, res=res) #shrunken results
resLFC
```
##***************Need to make these a variable that can be entered!!!!!**************
##Build a results table for contrasting Light Pulse 0 vs 30
```{r message=FALSE}
res0vs30 <- results(analysis, contrast=c("Light_Pulse","0_minutes","30_minutes"))
```

##Build a results table for contrasting Light Pulse 0 vs 60
```{r message=FALSE}
res0vs60 <- results(analysis, contrast=c("Light_Pulse","0_minutes","60_minutes"))
```

##Build a results table for contrasting Light Pulse 30 vs 60
```{r message=FALSE}
res30vs60 <- results(analysis, contrast=c("Light_Pulse","30_minutes","60_minutes"))
```

##Log fold change 0 vs 30
```{r include=FALSE, echo=TRUE}
resultsNames(analysis)
resLFC0vs30 <- lfcShrink(analysis, contrast=c("Light_Pulse","0_minutes","30_minutes"))
```

##Log fold change 0 vs 60
```{r include=FALSE, echo=TRUE}
resultsNames(analysis)
resLFC0vs60 <- lfcShrink(analysis, contrast=c("Light_Pulse","0_minutes","60_minutes"))
```

##Log fold change30 vs 60
```{r include=FALSE, echo=TRUE}
resultsNames(analysis)
resLFC30vs60 <- lfcShrink(analysis, contrast=c("Light_Pulse","30_minutes","60_minutes"))
```

##Order by p-value 0 vs 30
```{r include=FALSE, echo=TRUE}
res0vs30Ordered <- res0vs30[order(res0vs30$pvalue),]
```

##Order by p-value 0 vs 36
```{r include=FALSE, echo=TRUE}
res0vs60Ordered <- res0vs60[order(res0vs60$pvalue),]
```

##Order by p-value 0 vs 30
```{r include=FALSE, echo=TRUE}
resOrdered30vs60 <- res30vs60[order(res30vs60$pvalue),]
```

##Plot data
```{r message=FALSE}
plotMA(res0vs30, ylim=c(-2,2))
plotMA(res0vs60, ylim=c(-2,2))
plotMA(res30vs60, ylim=c(-2,2))
```

#Plot shrunken data
```{r message=FALSE}
plotMA(resLFC0vs30, ylim=c(-2,2))
plotMA(resLFC0vs60, ylim=c(-2,2))
plotMA(resLFC30vs60, ylim=c(-2,2))
```

##Find a gene, plot by light pulse length
```{r warning=FALSE, message=FALSE}
library(dplyr)
library(stringr)
mygene <- anno_table %>% filter(str_detect(SYMBOL, genetocheck))
mygene1 <- anno_table %>% filter(str_detect(GENENAME, genetocheck))
mygene2 <- rbind(mygene,mygene1)
mygene3 <- filter(anno_table,geneId==mygene2$geneId[1])
mypeak <- as.character(mygene3$peak[1])
print(mygene2)
if(nrow(mygene3) < 1){
  print("No genes")
} else{
  print(nrow(mygene2))
  print(mygene2$GENENAME[1])
  plotCounts(analysis, gene=mypeak, intgroup="Light_Pulse")
  plotCounts(analysis, gene=mypeak, intgroup="CT")
}
```

##Plot gene with lowest p-value
```{r warning=FALSE}
bestpeak<-rownames(res)[which.min(res$padj)]
bestgene <-filter(anno_table,peak==bestpeak)
g1 <- bestgene$geneId
g2 <- bestgene$GENENAME
plotCounts(analysis, gene=which.min(res$padj), intgroup="CT")
print(g2)
```

##Combining Conditions
```{r include=FALSE}
analysis$group <- factor(paste0(analysis$Light_Pulse, analysis$CT))
group <- design(analysis) <- ~ group
analysis_multi <- DESeq(analysis)
```

#Contrast CT16 LP0 mins vs CT16 LP30 mins, this can be changed to contrast LP 60 or CT
```{r}
resultsNames_multi <- resultsNames(analysis_multi)
results_multi <- results(analysis_multi, contrast=c("group", "30_minutesCT16", "0_minutesCT16"))
#results_multi_ordered <- results_multi[order(results_multi$pvalue),]
```

#Plot with gene of lowest p-value
```{r}
bestpeak2<-rownames(results_multi)[which.min(results_multi$pvalue)]
bestgene <-filter(anno_table,peak==bestpeak)
bestgene2 <-filter(anno_table,peak==bestpeak2)
print(bestgene2$GENENAME)
plotCounts(analysis_multi, gene=which.min(results_multi$pvalue), intgroup="Light_Pulse")
plotCounts(analysis_multi, gene=which.min(results_multi$pvalue), intgroup="CT")
```


##Data quality assessment
##Heatmap of the count matrix
##Set up the heatmap
```{r}
library("pheatmap")
select <- order(rowMeans(counts(analysis,normalized=TRUE)),
                decreasing=TRUE)[1:20]
df <- as.data.frame(colData(analysis)[,c("Light_Pulse","CT")])
ntd <- normTransform(analysis)
```

##Construct heatmap for the log2(n+1) data
```{r}
pheatmap(assay(ntd)[select,], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=df)
```


```{r}
rld <- rlog(analysis, blind=FALSE)
pheatmap(assay(rld)[select,], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=df)
```

##Heatmap of the sample-to-sample distances
```{r}
sampleDists <- dist(t(assay(rld)))
library("RColorBrewer")
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- paste(rld$Light_Pulse, rld$CT, sep="-")
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors)
```


##Principal component plot of the samples with Ggplot
```{r}
library(ggplot2)
pcaData <- plotPCA(rld, intgroup=c("Light_Pulse", "CT"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color=Light_Pulse, shape=CT)) +
  geom_point(size=3) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed()
```

#Boxplot if lots of ouliers
#```{r}
#par(mar=c(8,5,2,2))
#boxplot(log10(assays(analysis)[["cooks"]]), range=0, las=2)
#```

#```{r}
#plotDispEsts(analysis)
#```

