---
############################################################################################
#Work in progress!!									   #
#This code will be integrated into the pipeline to allow a user to normalise their counts  #
#table using a house keeping gene list							   #
############################################################################################
---


---
title: "RUVseq"
author: "Steve W."
date: "December 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load RUVseq, load in count file as matrix, set up experimental parametres as X (by CT only used for colouring), create expression set as 'set'
```{r include=FALSE}
library(RUVSeq)
library(stringr)
myfile <- "/ifs/projects/proj078/1_atac_seq/6_differential_analysis/2_paired_end/Peak_counts_copy.csv"
counts <- as.matrix(read.csv(myfile,sep=",",row.names="Peaks"))
x <- as.factor(rep(c(rep("CT8",10), rep("CT16", 12))))
set <- newSeqExpressionSet(as.matrix(counts),
phenoData = data.frame(x, row.names=colnames(counts)))
```

##Load ChIPseeker for peak annotation, load in macs2 peak file, annotate peaks
```{r}
library(ChIPseeker)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
peakfile <- "/ifs/projects/proj078/1_atac_seq/6_differential_analysis/2_paired_end/fully_merged_downsampled_pooled.macs2_peaks.narrowPeak"
peak <- readPeakFile(peakfile,header=FALSE)
peakAnno <- annotatePeak(peak, tssRegion=c(-3000, 3000),
                         TxDb=txdb, annoDb="org.Mm.eg.db")
anno_table <- as.data.frame(peakAnno)#        #Create data frame from gene annotations
colnames(anno_table)[6] <- "peak"#             #Name column 6 peak
counts_genes <- counts                        # set up new counts table for genes not peaks
row.names(counts_genes) <- anno_table[,23]    # set row names as genes rather than peaks
merged_gc <- aggregate(counts_genes,by=list(rownames(counts_genes)),FUN = "sum")  #Merge all peaks for each gene
row.names(merged_gc) <- merged_gc[,1]       #Set gene Ids as row names
merged_gc[,1] <- NULL#                     #remove gene ID column
set2 <- newSeqExpressionSet(as.matrix(merged_gc),#      #Create new gene expression set with gene counts table
phenoData = data.frame(x, row.names=colnames(merged_gc)))
```

#Carry out RLE and PCA on data before any normalisation (For gene counts table)
```{r}
library(RColorBrewer)
colors <- brewer.pal(3, "Set2")
plotRLE(set2, outline=FALSE, ylim=c(-4, 4), col=colors[x])
plotPCA(set2, col=colors[x], cex=1.2)
```

#Normalise by upper quartile and plot RLE/PCA (For Gene counts table)
```{r}
set2 <- betweenLaneNormalization(set2, which="upper")
plotRLE(set2, outline=FALSE, ylim=c(-4, 4), col=colors[x])
plotPCA(set2, col=colors[x], cex=1.2)
```

##Test gene counts table against emperical controls
```{r}
design <- model.matrix(~x, data=pData(set))
y <- DGEList(counts=counts(set), group=x)
y <- calcNormFactors(y, method="upperquartile")
y <- estimateGLMCommonDisp(y, design)
y <- estimateGLMTagwiseDisp(y, design)
fit <- glmFit(y, design)
lrt <- glmLRT(fit, coef=2)
top <- topTags(lrt, n=nrow(set))$table
empirical <- rownames(set2)[which(!(rownames
                                   (set2) %in% rownames(top)[1:5000]))]
set3 <- RUVg(set2, empirical, k=1)
```

#Plot PCA/RLE against gene counts table normalised by empirical controls
```{r}
plotRLE(set3, outline=FALSE, ylim=c(-4, 4), col=colors[x])
plotPCA(set3, col=colors[x], cex=1.2)
```

##Load in light pulse data from sik paper and time of day data from  Olive et al 2015. Find genes not induced by light from aarti data, ie house keeping genes
```{r}
myfile2 <- "/ifs/projects/proj078/1_atac_seq/8_RUVseq/aarti_sik1.csv"
myfile3 <- "/ifs/projects/proj078/1_atac_seq/8_RUVseq/time_of_day.csv"
aarti_rna <- as.data.frame(read.csv(myfile2,sep=",", stringsAsFactors=FALSE))
timeofday <- as.data.frame(read.csv(myfile3,sep=",",header=TRUE, stringsAsFactors=FALSE))
aarti_rna$LP30induction <- (aarti_rna$LP30 - aarti_rna$Sham)
aarti_rna$LP60induction <- (aarti_rna$LP60 - aarti_rna$Sham)
```

##Filter out genes that are expressed but whose expression does not change by light or time of day
```{r}
library(dplyr)
hk_time <- filter(timeofday, Dynamic_range < 2, Expression_level > 100, DESeq2.padj < 0.05)
hk_time_genes <- hk_time$Gene_ID # Create a list of the genes
hk_light <- filter(aarti_rna, LP30induction < 0.25, LP30induction > -0.25, LP60induction < 0.25, LP60induction > -0.25)
hk_light_genes <- hk_light$Gene_Symbol
genes_in_both <- hk_light_genes[which(hk_light_genes %in% hk_time_genes)] #Create a list of genes in both
```

##Venn plot to check gene lists
```{r}
require("gplots")
venn(list(first.vector = hk_time_genes, second.vector = hk_light_genes))
```

##Create count tables from hk gene lists, data extracted from set2
```{r}
hk_light_counts <- rownames(set2)[which(!(hk_light_genes %in% unlist(rownames(merged_gc))))]
hk_time_counts <- rownames(set2)[which(!(hk_time_genes %in% unlist(rownames(merged_gc))))]
hk_both_counts <- rownames(set2)[which(!(genes_in_both %in% unlist(rownames(merged_gc))))]
```

##Create an expression set for each of the house keeping gene sets
```{r}
set4 <- RUVg(set2, hk_light_counts, k=1)
set5 <- RUVg(set2, hk_time_counts, k=1)
set6 <- RUVg(set2, hk_both_counts, k=1)
```

##Plot PCA/RLE for each expression set normalised by hk genes
```{r}
plotRLE(set4, outline=FALSE, ylim=c(-4, 4), col=colors[x])
plotPCA(set4, col=colors[x], cex=1.2)
plotRLE(set5, outline=FALSE, ylim=c(-4, 4), col=colors[x])
plotPCA(set5, col=colors[x], cex=1.2)
plotRLE(set6, outline=FALSE, ylim=c(-4, 4), col=colors[x])
plotPCA(set6, col=colors[x], cex=1.2)
```

##This code creates a new counts table for genes with peaks near their promoters
```{r}
anno_tableprom <- as.data.frame(peakAnno) #Make a new copy of the annotation table
colnames(anno_tableprom)[6] <- "peak"#             #Name column 6 of new table peak
counts_prom <- counts                        # set up new counts table for promoters not peaks (or genes)
row.names(counts_prom) <- anno_tableprom[,23]    # set row names as genes (nearest promoters) rather than peaks
```

```{r}
anno_tableprom <- filter(anno_tableprom, grepl("Promoter",annotation))  #Remove all genes that do not have peaks near promoters
counts_prom <-counts_prom[(which(rownames(counts_prom) %in% anno_tableprom[,23])),] # Remove all genes that are not in the promoter annotation
```

```{r}
#table (The one giving the genes with peaks near promoters)
merged_pc <- aggregate(counts_prom,by=list(rownames(counts_prom)),FUN = "sum")  #Merge all peaks for each gene in new promoter counts table
row.names(merged_pc) <- merged_pc[,1]       #Set gene Ids as row names
merged_pc[,1] <- NULL#                     #remove gene ID column
setprom <- newSeqExpressionSet(as.matrix(merged_pc),#      #Create new gene expression set with promoter gene counts table
phenoData = data.frame(x, row.names=colnames(merged_pc)))
```

```{r}
hk_light_counts <- rownames(counts_genes)[which(!(hk_light_genes %in% (rownames(counts_genes))))]  ##Which not working!!!
hk_time_counts <- rownames(set2)[which(!(hk_time_genes %in% unlist(rownames(merged_gc))))]
hk_both_counts <- rownames(set2)[which(!(genes_in_both %in% unlist(rownames(merged_gc))))]
```


##Make new counts tables for hk genes, but only for genes with peaks near promoters
```{r}
hk_light_counts_prom<- rownames(counts_prom)[which(!(hk_light_genes %in% rownames(counts_prom)))]
hk_time_counts_prom <- rownames(counts)[which(!(hk_time_genes %in% unlist(rownames(counts))))]
hk_both_counts_prom <- rownames(counts)[which(!(genes_in_both %in% unlist(rownames(counts))))]
```



```{r}
hk_light_counts_test<- hk_light_counts_prom[which(hk_light_counts_prom %in% hk_light_counts)]
```



